
---
title: "Statistical inference project"
author: "Zhenning Xu"
date: "June 9th, 2019"
output: 
  html_document:
    toc: true
    toc_float: true
---
---
  title: "cssTest"
    output:
    html_document:
      css: custom.css
      toc: yes
---
---
TOC {
  position: fixed;
  left: 0;
  top: 0;
  width: 200px;
  height: 100%;
  overflow:auto;
}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. I will set lambda = 0.2 for all of the simulations. I will investigate the distribution of averages of 40 exponentials. Note that I will need to do a thousand simulations.

## Simulations

```{r}
sims = 1000;
n = 40; ## number of distributions
lambda = 0.2; ## number of simulations
means <- vector("numeric")
means_sum <- vector("numeric")
means_cum <- vector("numeric")

for (i in 1:sims) { means[i] <- mean(rexp(n, lambda))}
means_sum[1] <- means[1]
for (i in 2:sims) { means_sum[i] <- means_sum[i-1] + means[i] }
for (i in 1:sims) { means_cum[i] <- means_sum[i]/i }

```

## Sample Mean versus Theoretical Mean
```{r}
means_cum[sims]  #Sample mean
1/lambda  #Theoretical mean

```
We observed the distribution of means is centered in the theoretical mean.

## Sample variance versus Theoretical variance
```{r}

var(means)*n  #Sample variance
(1/lambda)^2  #Theoretical variance

```
We see that the theoretical variance and the variance of the means are comparable.

## Visualize Sample Mean vs Theoretical  Mean
```{r}
library(ggplot2)
g <- ggplot(data.frame(x = 1:sims, y = means_cum), aes(x = x, y = y))
g <- g + geom_hline(yintercept = 0) + geom_line(size = 2)
g <- g + geom_abline(intercept = 1 / lambda, slope = 0, color = "red", size = 1)
g <- g + scale_y_continuous(breaks=c(4.50, 4.75, 5.00, 5.25, 5.50, 5.75), limits=c(4.25, 6))
g <- g + labs(title="Sample Mean vs Theoretical  Mean")
g <- g + labs(x = "Simulations", y = "Sample Mean")
print(g)

```

We see that Sample Mean vs Theoretical  Mean are comparable 

## Normal Distribution


```{r}
library(ggplot2)
g <- ggplot(data.frame(x = means), aes(x = x))
g <- g + geom_histogram(position="identity", fill="green", color="red", alpha=0.2,binwidth=0.5, aes(y= ..density..))
g <- g + stat_function(fun = dnorm, colour = "red", args=list(mean=5))
g <- g + scale_x_continuous(breaks=c(1, 2, 3, 4, 5, 6, 7, 8, 9), limits=c(1, 9))
g <- g + scale_y_continuous(breaks=c()) 
g <- g + theme(plot.title = element_text(size=12, face="bold", vjust=2, hjust=0.5))
g <- g + labs(title="Distribution of Samle Means vs Normal Distribution")
g <- g + labs(x = "Sample Mean", y = "Frequency")
print(g)  

```

The figure above also shows the density computed using the histogram and the normal density plotted with theoretical mean and variance values. 

Additionally, the q-q plot below indicates normality. The theoretical quantiles again match closely with the actual quantiles. Therefore, we conclude that the distribution is approximately normal.

```{r}
qqnorm(means, 
       main ="Q-Q plot of the standardized Sample Means")
qqline(means, 
       col = "3")
```







